---
title: "STAT 501 R HW1"
author: "McBeth Ahortor"
date: "9/4/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rmisc)
library(pastecs)
library(VIM)
```

## Problems Using R (Q.1.)

## Generating a random sample of size 24 from a normal distribution with mean = 21 and sd = 3 and then adding a 25th observation of 52.2

```{r, echo=T}
set.seed(3)
x <- rnorm(24, mean = 21, sd = 3)
x
y <- c(x,52.2)
y
# Concatenating 52.2 to the 24 random sample
hist(y, freq = FALSE, nclass = 20, col = "yellow", xlab = "Random numbers", main = "Histogram of 25 Random Numbers")

```

## a. Mean and standard deviation of the 25 random numbers


```{r, echo=T}
u <- mean(y)
u
#mean
s <- sd(y)
#Standard deviation
s
```

## b. Intervals 
For k=1,
```{r, echo=T}
k1 <- c(u-1*s,u+1*s)

k1

```
For k=2,
```{r, echo=T}
k2 <- c(u-2*s,u+2*s)
k2

```
For k=3,
```{r, echo=T}
k3 <- c(u-3*s,u+3*s)
k3

```
## C. Counting the number of measurements

```{r, echo=T}
t1 <-print( sum(u-1*s< y&y < u+1*s))
t2 <-print( sum(u-2*s< y&y < u+2*s))
t3 <-print( sum(u-3*s< y&y < u+3*s))

print("24 measurements fell in the intervals k1, k2 and k3")

```

For the empirical rule,

```{r, echo=T}
25*(0.68) # We expect 68% of the measurements to fall within the interval
25*(0.95) # We expect 95% of the measurements to fall within the interval
25*(1) # We expect almost all of the measurements to fall within the interval
```


## d. Range
Finding the range for the 25 random sample,
```{r, echo=T}
r <- range(y)
R <-diff(r)
R
print('The range = 36.19942')

```


Per empirical rule, the standard deviation can be approximated as range/4
```{r, echo=T}
sbar = R/4
sbar
print('The approximated standard error = 9.049856')

```


## e. Removing the outlier
```{r, echo=T}
Y <- y[-25]
Y

# Calculating the mean and standard deviation without the outlier;
ybar <-mean(Y)
S <- sd(Y)

print('The mean = 20.15815 and the standard deviation = 2.368674')

```
Yes, there is a wide difference in the two standard deviations and the two means. the disparity is as a result of the outlier.



```


























